---
title: "Welcome to the Geospatial Community"
image: turkeypin.svg

listing:
  - id: latest
    contents: posts
    sort: "date desc"
    type: default
    max-items: 1
  - id: all
    contents: posts
    sort: "date desc"
    type: table 
    fields: [image, date, title]
    image-height: 60px
  # - id: testimonials
  #   contents: testimonials
  #   sort: "date desc"
  #   type: grid
  #   max-items: 3
  #   grid-columns: 1

page-layout: article
title-block-banner: true

format:
  html:
    grid: 
      margin-width: 400px

---

```{=html}
<style>
#title-block-header{
  display: none;
}
#all-posts h2 {
  margin-top: 0px;
}
div.quarto-post {
  margin-bottom: 0px;
}
.tui-view-7 {
height: 100% !important;
}
.tui-full-calendar-month {
    height: 75%;
    min-height: 500px;
}
.tui-full-calendar-popup-detail .tui-full-calendar-popup-container {
    width: 330px;
}
.card-img{
  width: 100px;
}
img {
    width: 100px;
    height: 100% !important;
    max-height: 100px !important;
}

.column-margin.column-container>*:not(.collapse) {
    padding-top: 14em;
}

@media (max-width: 767.98px){
  #events-calendar{
    display: none;
  }
}
</style>

<!-- html code containing the turkey pin and page title -->

<div style="padding-bottom: 2em;">
  <div style="display: inline-block;width: 89px;">
    <img src="turkeypin.svg" style="width: 4em; vertical-align: -50px;" class="GeoCommunity Turkey Pin Logo">
  </div>

  <div style="display: inline-block;width: 70%;">
    <h1 class="title">Geospatial Community</h1>
  </div>
</div>
```

::: column-margin
## Events Calendar
```{r calendar}
#| include: false

library(googlesheets4)
#gs4_auth(cache = ".secrets", email = "itsuqnick@gmail.com")
gs4_deauth()
cal_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HHxs4SsFO1ImH5XbMgTwK2QrheN8d8N3_9P5L0QiPUg/edit#gid=308321545", sheet = "confirmed")
cal_prop <- read_sheet("https://docs.google.com/spreadsheets/d/1HHxs4SsFO1ImH5XbMgTwK2QrheN8d8N3_9P5L0QiPUg/edit#gid=308321545", sheet = "cal_properties")

#install.packages("toastui")
library(toastui)
```

```{r }
#| echo: false
#| warning: false
#| error: false

calendar(cal_data, navigation = TRUE, defaultDate = Sys.Date(), height = "600") %>% 
  cal_month_options(
    startDayOfWeek  = 1, 
    workweek = TRUE
  ) %>% 
  cal_props(cal_prop)
```
:::

Our mission is to build a **supportive** and **inclusive** community where students, researchers, and professionals can **grow their geospatial skills together**. 🌱


::: {.callout-note icon="false" collapse="true"}
# Want to **join** our Geospatial Community workshops?

<br> We run **monthly skill sharing tutorial sessions** where a member of our community teaches us something they have learned along the way.

Fill out [this form](https://docs.google.com/forms/d/e/1FAIpQLSeXfv6EPzm6ur5e9IrDPK07e5N1y2xxYNQlBe_xutTDu-ajBw/viewform) to join and get the workshop link:

```{=html}
<iframe width="100%" height="400px" src="https://docs.google.com/forms/d/e/1FAIpQLSeXfv6EPzm6ur5e9IrDPK07e5N1y2xxYNQlBe_xutTDu-ajBw/viewform" title="R for mapping and more"></iframe>
```
:::

:::{.callout-important collapse="true"}
## **Petition** to increase Australian PhD Stipends

Please consider signing this [e-petition](https://www.aph.gov.au/e-petitions/petition/EN5085) to increase PhD stipends to minimum wage. Petition reason from the online petition is copied below.

Petition Reason

PhD students contribute enormously to Australia’s research every year, with their time accounting for 56% the total research output. They are the future professors and industry professionals of this country, responsible for driving innovation and ensuring strong international collaborations and competition. Their contribution is often rewarded with a stipend intended to cover the cost of living, with a minimum rate set by the Australian government. Currently, this scholarship rate is $29,863 per year for PhD students to live on while usually working more than 40 hrs per week on their research. Not only is this below minimum wage, but it is even forcing students to live below the poverty line. With this treatment, students are incentivised to accept positions abroad, reject higher degrees by research, or quit midway through their degree due to financial distress. Australia is actively discouraging higher education and slowing its economic growth and productivity by underpaying PhD students. We believe the government should support higher education and ensure a competitive standing for Australia. The way to do this is by setting a fair, liveable rate for PhD scholarships, which is at least equivalent to the minimum wage.
:::

## Upcoming Workshop {style="padding-top: 5px;"}

### Interactive Map with Python + Folium

#### May 18th, 3-5pm AEST on Teams

[Dr Tony Howes](https://chemeng.uq.edu.au/profile/2223/tony-howes) (UQ, School of Chemical Engineering) will be sharing how Python and Folium (a wrapper for Python of the JavaScript Leaflet library) work together to create an interactive map.

The  data will be from the Brisbane Dawn Nav, May 5th MapRun. Workshop materials will be in a Jupyter notebook and downloading [Anaconda](https://www.anaconda.com/) is recommended if you wish to follow along.


## Last Post

::: {#latest}

:::

## All Posts

::: {#all}
:::
