---
title: "Welcome to the Geospatial Community"
image: turkeypin.svg

listing:
  - id: latest
    contents: posts
    sort: "date desc"
    type: default
    max-items: 1
  - id: all
    contents: posts
    sort: "date desc"
    type: table 
    fields: [image, date, title]
    image-height: 60px
  # - id: testimonials
  #   contents: testimonials
  #   sort: "date desc"
  #   type: grid
  #   max-items: 3
  #   grid-columns: 1

page-layout: article
title-block-banner: true

format:
  html:
    grid: 
      margin-width: 400px
---

```{=html}
<style>
#title-block-header{
  display: none;
}
#all-posts h2 {
  margin-top: 0px;
}
div.quarto-post {
  margin-bottom: 0px;
}
.tui-view-7 {
height: 100% !important;
}
.tui-full-calendar-month {
    height: 75%;
    min-height: 500px;
}
.tui-full-calendar-popup-detail .tui-full-calendar-popup-container {
    width: 330px;
}
.card-img{
  width: 100px;
}
img {
    width: 100px;
    height: 100% !important;
    max-height: 100px !important;
}

.column-margin.column-container>*:not(.collapse) {
    padding-top: 14em;
}

@media (max-width: 767.98px){
  #events-calendar{
    display: none;
  }
}
</style>

<!-- html code containing the turkey pin and page title -->

<div style="padding-bottom: 2em;">
  <div style="display: inline-block;width: 89px;">
    <img src="turkeypin.svg" style="width: 4em; vertical-align: -50px;" class="GeoCommunity Turkey Pin Logo">
  </div>

  <div style="display: inline-block;width: 70%;">
    <h1 class="title">Geospatial Community</h1>
  </div>
</div>
```
::: column-margin
## Events Calendar

```{r calendar}
#| include: false

library(googlesheets4)
#gs4_auth(cache = ".secrets", email = "itsuqnick@gmail.com")
gs4_deauth()
cal_data <- read_sheet("https://docs.google.com/spreadsheets/d/1HHxs4SsFO1ImH5XbMgTwK2QrheN8d8N3_9P5L0QiPUg/edit#gid=308321545", sheet = "confirmed")
cal_prop <- read_sheet("https://docs.google.com/spreadsheets/d/1HHxs4SsFO1ImH5XbMgTwK2QrheN8d8N3_9P5L0QiPUg/edit#gid=308321545", sheet = "cal_properties")

#install.packages("toastui")
library(toastui)
```

```{r }
#| echo: false
#| warning: false
#| error: false

calendar(cal_data, navigation = TRUE, defaultDate = Sys.Date(), height = "600") %>% 
  cal_month_options(
    startDayOfWeek  = 1, 
    workweek = TRUE
  ) %>% 
  cal_props(cal_prop)
```
:::

Our mission is to build a **supportive** and **inclusive** community where students, researchers, and professionals can **grow their geospatial skills together**. 🌱

::: {.callout-note icon="false" collapse="true"}
# Want to **join** our Geospatial Community workshops?

<br> We run **monthly skill sharing tutorial sessions** where a member of our community teaches us something they have learned along the way.

Fill out [this form](https://docs.google.com/forms/d/e/1FAIpQLSeXfv6EPzm6ur5e9IrDPK07e5N1y2xxYNQlBe_xutTDu-ajBw/viewform) to join and get the workshop link:

```{=html}
<iframe width="100%" height="400px" src="https://docs.google.com/forms/d/e/1FAIpQLSeXfv6EPzm6ur5e9IrDPK07e5N1y2xxYNQlBe_xutTDu-ajBw/viewform" title="R for mapping and more"></iframe>
```
:::

## Upcoming Workshop {style="padding-top: 5px;"}

### More spatial data in Python!

#### June 15th, 2-4pm AEST on Teams

Join us for an exciting session where [Dr César Herrara Acosta](https://experts.griffith.edu.au/30877-c%C3%A9sar-herrera-acosta) (Griffith University) will take you on a new dive into the dynamic world of the Python programming language ecosystem! Building upon previous sessions, César will showcase Python libraries for (1) analysis and visualization of spatial data (e.g. geopandas, datashader, holoviews) and (2) basic computer vision algorithms (OpenCV). César will also make use of the incredibly versatile Jupyter Notebooks and Jupyter Lab frameworks, which empowers users with seamless workflows for code development, data exploration, and interactive presentations.

## Last Post

::: {#latest}
:::

## All Posts

::: {#all}
:::
